---
title: "Bellabeat Case Study"
author: Sharon Viers
output: html_document
date: "2025-03-18"
---

### Statement of Business task:
#### Bellabeat, a high-tech manufacturer of health-focused products for women
#### wants to identify new growth opportunities for their smart device. The analysis  
#### seeks to answer the following stakeholder questions:
#### - What are some of the trends in smart device usage?
#### - How could these trends apply to Bellabeat customers?
#### - How could these trends help influence Bellabeat marketing straegy?

### Data source
#### FitBit Fitness Tracker Data - https://www.kaggle.com/datasets/arashnic/fitbit


### Install packages
```{r install packages}
install.packages("tidyverse")
install.packages("skimr")
install.packages("janitor")
install.packages("esquisse")
library(tidyverse)
library(skimr)
library(janitor)
library(esquisse)
```

### Create dataframes
``` {r create datatframes}
library(readr)
daily_activity <- read_csv("CaseStudy/dailyActivity_merged.csv")
sleep_day <- read_csv("CaseStudy/sleepDay_merged.csv")
daily_intensity <- read_csv("CaseStudy/dailyIntensities_merged.csv")
daily_steps <- read_csv("CaseStudy/dailySteps_merged.csv")
heart_rate <- read_csv("CaseStudy/heartrate_seconds_merged.csv")
```
### Add NA to blank cells
```{r}
daily_activity[daily_activity == ""] <- NA
sleep_day[sleep_day == ""] <- NA
daily_intensity[daily_intensity == ""] <- NA
daily_steps[daily_steps == ""] <- NA
heart_rate[heart_rate == ""] <- NA
```

### Clean data by omitting NA
```{r}
daily_activity <- na.omit(daily_activity)
sleep_day <- na.omit(sleep_day)
daily_intensity <- na.omit(daily_intensity)
daily_steps <- na.omit(daily_steps)
heart_rate <- na.omit(heart_rate)
```

### Check columns and rows associated with data
``` {r columns and rows}
head(daily_activity)
head(sleep_day)
head(daily_intensity)
head(daily_steps)
head(heart_rate)

colnames(daily_activity)
colnames(sleep_day)
colnames(daily_intensity)
colnames(daily_steps)
colnames(heart_rate)
```


### Check for unique participants in all data by ID
``` {r unique participants}
n_distinct(daily_activity$Id)
n_distinct(daily_intensity$Id)
n_distinct(daily_steps$Id)
n_distinct(heart_rate$Id)
n_distinct(sleep_day$Id)
```
### Summary statistics
``` {r daily activity statistics}
daily_activity %>%
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()
```

```{r Summary statistics of daily intensities}
daily_intensity %>%
    select(SedentaryActiveDistance,
           LightActiveDistance,
           ModeratelyActiveDistance,
           VeryActiveDistance) %>%
  summary()
```

```{r statistics of daily steps}
daily_steps %>%
  select(StepTotal)%>%
  summary()
```

```{r statistics of heart rate}
heart_rate %>%
  select(Value)%>%
  summary()

```

```{r statistics of sleep day}
sleep_day%>%
  select(TotalMinutesAsleep,
         TotalTimeInBed)%>%
  summary()

```

``` {r activity}
ggplot(daily_activity) +
 aes(x = TotalDistance, y = Calories, colour = Calories) +
 geom_jitter() +
 geom_jitter(aes(x = TotalDistance, 
 y = Calories)) +
 scale_color_viridis_c(option = "viridis", direction = 1) +
  labs(
    x = "Total Distance",
    y = "Calories",
    title = "Distance in Miles vs. Calories Burned"
  ) 
 theme_minimal()
```

```{r steps vs distance}
ggplot(daily_activity) +
  aes(
    x = TotalSteps,
    y = TotalDistance,
    colour = TotalDistance
  ) +
  geom_point(shape = "diamond") +
  scale_color_viridis_c(option = "plasma", direction = 1) +
  labs(
    x = "Total Steps",
    y = "Total Distance",
    title = "Steps per Day vs. Distance in Miles"
  ) +
  theme_minimal()
```

```{r very active users}
ggplot(daily_intensity) +
  aes(x = VeryActiveMinutes, y = VeryActiveDistance) +
  geom_step(linewidth = 1.75, colour = "#FF8C00") +
  labs(
    title = "Very Active Users",
    subtitle = "Minutes vs. Distance"
  ) +
  theme_minimal()
```

```{r steps vs. sedentary minutes}
ggplot(daily_activity) +
  aes(x = TotalSteps, y = SedentaryMinutes) +
  geom_smooth(span = 0.31, se = TRUE, colour = "#319827") +
  labs(title = "Total Steps vs Sedentary Minutes") +
  theme_minimal()
```

```
```{r minutes asleep vs time in bed}
ggplot(sleep_day) +
  aes(
    x = TotalMinutesAsleep,
    y = TotalTimeInBed,
    colour = TotalMinutesAsleep
  ) +
  geom_point(size = 2.9, shape = "diamond") +
  scale_color_viridis_c(option = "inferno", direction = 1) +
  labs(title = "Minutes Asleep vs. Time in Bed") +
  theme_minimal()
```

``` {r write new combined_data file}
library(tidyverse)
sleep_day <- read_csv("CaseStudy/sleepDay_merged.csv")
daily_activity <- read_csv("CaseStudy/dailyActivity_merged.csv")
combined_data <- full_join(sleep_day, daily_activity, by = "Id")
write_csv(combined_data, "CaseStudy/combined_data.csv")
list.files("CaseStudy")
```

```{r }
ggplot(combined_data) +
  aes(
    x = TotalMinutesAsleep,
    y = TotalSteps,
    colour = Calories
  ) +
  geom_jitter() +
  scale_color_distiller(palette = "Paired", direction = 1) +
  labs(
  x = ("Total Minutes Asleep"),
  y = ("Total Steps"),
  title = ("Total Minutes Asleep vs. Total Steps")
)
  theme_minimal()
```


